---
title: "w8b_lecture"
author: "Micaela Chapuis"
date: "2024-10-17"
output: html_document
---
## Advanced plotting

{patchwork} - bringing plots together
{ggrepel} - repel your labels on your plots
{gganimate} - add some animations to your plots
{magick} - photo processing


```{r}
library(tidyverse)
library(here)
library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)
```


### Patchwork
Easily bring your plots together.  

Let's start with two simple plots from the Palmer penguin dataset
```{r}
# plot 1
p1<-penguins %>%
  ggplot(aes(x = body_mass_g, 
             y = bill_length_mm, 
             color = species))+
  geom_point()
p1
```

```{r}
# plot 2
p2<-penguins %>%
  ggplot(aes(x = sex, 
             y = body_mass_g, 
             color = species))+
  geom_jitter(width = 0.2)
p2
```

Bring the plots together using simple operations
```{r}
p1 + p2 +
  plot_layout(guides = "collect") + # collects all legends and puts only one on the left
  plot_annotation(tag_levels = "A") # if I put lowercase a it would use lower case a, b, etc
```

Put one plot on top of the other
```{r}
p1 / p2 +
  plot_layout(guides = "collect") + # collects all legends and puts only one on the left
  plot_annotation(tag_levels = "A")
```

### ggrepel
Use the mtcars dataset
```{r}
view(mtcars)
```

```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) + # gets the row names and makes labels out of them
  geom_text() + # creates a text label with the car name at the corresponding x,y coordinate
  geom_point(color = 'red')
```
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) + # gets the row names and makes labels out of them
  geom_text_repel() + # makes the labels move away from the points so you can read them
  geom_point(color = 'red')
```
We can also use the label function instead of text
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) +
  geom_label_repel() + # creates a label and repels them
  geom_point(color = 'red')
```

### gganimate

```{r}

penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states( # adding transition
     year, # what are we animating by
     transition_length = 2, # the relative length of the transition (how long it takes to go from one year to the next)
     state_length = 1) + # the length of the pause between transitions 
  ease_aes("bounce-in-out") + # specify how it changes from one state to the next 
  labs(title = "Year: {closest_state}") +
  anim_save(here("Week8","Output", "Lecture", "mypenguingif.gif"))
```


### magick
For advanced image processing


Get a penguin pic
```{r}
penguin <- image_read("https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcRS_TdFcDG2LBfSRY5ld6swoXVmy6xrZ_FQDQ&s") # read in file from online (or can read it in from our folders using here)
penguin
```


Now let's put it on a penguin plot:  
**To put it on a plot you first need to save your plot as an image.**
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point()

ggsave(here("Week8","Output", "Lecture", "penguinplot.png"))
```

Then read in the plot as a magick image and make a composite plot
```{r}
penplot<-image_read(here("Week8","Output", "Lecture", "penguinplot.png"))
out <- image_composite(penplot, penguin, offset = "+70+30")
out
```




We can do this with gifs too!
```{r}
pengif<-image_read("https://media1.tenor.com/m/h5xD3gOTigUAAAAC/pesto-pesto-penguin.gif")
outgif <- image_composite(penplot, pengif, gravity = "center") # gravity says where we will put it
animation <- image_animate(outgif, fps = 10, optimize = TRUE) # because we have a gift it's an animation so we have to do this and give it fps
animation

# we can save this with anim_save
```

