---
title: "w9a_lecture"
author: "Micaela Chapuis"
date: "2024-10-22"
output: html_document
---

## Why Functions 

- whenever you want to do a task more than once
- keep your script clean
- avoid mistakes with copy and paste
- minimize places where you need to update your code

## Load Libraries 
```{r}
library(tidyverse)
library(palmerpenguins)
library(PNWColors)
```

## Making a dataframe of random numbers
```{r}
df <- tibble(
  a = rnorm(10), # draws 10 random values from a normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)
```

## Example 
We want to rescale data in multiple columns (value - min/(max - min))  

### Rescale all data individually:
```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(b, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))

df
```

### Instead, write a function

```{r}
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
```

```{r}
df %>%
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))
```

There are four key steps to creating a new function:

1. You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.
2. You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).
3. You place the code you have developed in body of the function, a { block that immediately follows function(...).
4. return tells us what values we want returned when you run the function



## Fahrenheit to Celsius
```{r}
# temp_C <- (temp_F - 32) * 5/9 # first figure out calculation for F to C

fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5/9
  return(temp_C)
}
```

```{r}
fahrenheit_to_celsius(32)
```


### Think, pair, share
```{r}
celsius_to_kelvin <- function(temp_C) {
  temp_K <- temp_C + 273.15
  return(temp_K)
}
```

```{r}
celsius_to_kelvin(-273.15)
```


## Making plots into a function

Let's say you have a specifically formatted plot that you like and that you plan to use over and over again. By turning it into a function, you only have to code it once.

```{r}
pal<-pnw_palette("Lake", 3, type = "discrete") # my color palette, specifying we only want 3 discrete colors
```

```{r}
ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal) +   # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()
```

Now make it a function  
{rlang} uses what is literally called a"curly-curly" {{}} within the body of the function to help us assign variables that are column names in dataframes. --> so it knows to search for those things within the dataframe
```{r}
myplot <- function(data, x, y) {
  pal<-pnw_palette("Lake", 3, type = "discrete") # my color palette, specifying we only want 3 discrete colors

  ggplot(data, aes(x = {{x}}, y = {{y}}, color = island)) +
  geom_point()+
  geom_smooth(method = "lm") + # add a linear model
  scale_color_manual("Island", values=pal) + # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()
}
```

Now test:  
```{r}
myplot(data = penguins, x = body_mass_g, y = bill_length_mm)
```

Test with other variables:  
```{r}
myplot(data = penguins, x = body_mass_g, y = flipper_length_mm)
```

### Adding Defaults
Let's say you want to create a default for the function to always default to the penguins dataset. You can set those directly in the function.
```{r}
myplot2 <- function(data = penguins, x, y) {
  pal<-pnw_palette("Lake", 3, type = "discrete") # my color palette, specifying we only want 3 discrete colors

  ggplot(data, aes(x = {{x}}, y = {{y}}, color = island)) +
  geom_point()+
  geom_smooth(method = "lm") + # add a linear model
  scale_color_manual("Island", values=pal) + # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()
}
```


Now we can write a function and only specify x and y
```{r}
myplot2(x = body_mass_g, y = flipper_length_mm)
```

### Layering the plot
Can add stuff with + like a regular ggplot
```{r}
myplot2(x = body_mass_g, y = flipper_length_mm) +
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")
```

### If-else statements
Imagine you want a variable to be equal to a certain value if a condition is met. This is a typical problem that requires the if ... else ... construct. For instance:
```{r}
a <- 4 
b <- 5
```

Suppose that if a > b then f should be = to 20, else f should be equal to 10. Using if/else we:
```{r}
if (a > b) { # my question
  f <- 20 # if it is true give me answer 1
    } else { # else give me answer 2
  f <- 10
}

f
```

#### Back to plotting
Using if...else... we can make our function even more flexible. Let's say we want the option of adding the geom_smooth lines. We can create a variable that if set to TRUE add the geom_smooth, otherwise print without.

First add a new argument for lines and make the default TRUE for ease

```{r}
myplot<-function(data = penguins, x, y, lines=TRUE ) { # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 

if(lines==TRUE) {
  ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
} else { 
  ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+ # no geom_smooth
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
  } # for the else statement
} # for the main function
```

Testing  
With lines:
```{r}
myplot(x = body_mass_g, y = flipper_length_mm) # because the default is lines = TRUE
```

Without lines:
```{r}
myplot(x = body_mass_g, y = flipper_length_mm, lines = FALSE)
```

