---
title: "w7a_lecture"
author: "Micaela Chapuis"
date: "2024-10-08"
output: html_document
---

## Notes

**Some Helpful Packages**  
(other than 'maps', 'mapproj' and 'mapdata')  
- sf (This is for the serious GIS people. If you want to do actual spatial analyses this is the package to use)  
- ggmap (Uses the google maps API to make maps. Really easy to use, but requires a credit card for the API...)  
- tmap (static and interactive maps)  
- leaflet (interactive html maps-- Awesome to use with rmakrdown files)  

## Code

### Loading Libraries
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapproj)
library(mapdata)
```

### Loading in data
```{r}
# Read in data on population in California by county
popdata<-read_csv(here("Week7","Data","CApopdata.csv"))
#read in data on number of seastars at different field sites
stars<-read_csv(here("Week7","Data","stars.csv"))
```


### 'Maps' Package
The {maps} package is a combination of functions that pair well with ggplot and base layers for maps (i.e. polygons of regions of interest).  
- map_data("location") is the function used to pull out whatever base layer that you want

This package doesn't get updated often so shouldn't be used for cases when limits really matter.  
It's also US centric, so we can only look at states, counties, etc for the US.  
Also this data is very coarse, so for really small islands (like Mo'orea in the world map) it will give a square-ish polygon.  

We can extract polygons for different countries:  
- Entire world:
```{r}
world <- map_data("world") # get data for the entire world
head(world)
```
- USA (Hawaii is left out):
```{r}
usa <- map_data("usa")
head(usa)
```

We can also extract data at different resolutions. For example, US states:
```{r}
states <- map_data("state")
head(states)
```

```{r}
counties <- map_data("county")
head(counties)
```

#### Data Structure
- **long** is longitude. Things to the west of the prime meridian are negative.
- **lat** is latitude.
- **order**. This just shows in which order ggplot should “connect the dots”
- **region and subregion** tell what region or subregion a set of points surrounds.
- **group.** This is very important! ggplot2’s functions can take a group argument which controls (amongst other things) whether adjacent points should be connected by lines. If they are in the same group, then they get connected, but if they are in different groups then they don’t. Essentially, having to points in different groups means that ggplot “lifts the pen” when going between them.

#### Making a map of the world
```{r}
ggplot() + 
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group,
                   fill = region), # fill by country
               color = "black") + # add color to the lines and fill
  guides(fill = FALSE) + # don't want the legend
  theme_minimal() +
  theme(panel.background = element_rect(fill = "lightblue")) +
  coord_map(projection = "mercator", # specify that we want to use the Mercator projection
            xlim = c(-180,180)) # set x limits
```

Alternative projection
```{r}
ggplot() + 
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group,
                   fill = region), # fill by country
               color = "black") + # add color to the lines and fill
  guides(fill = FALSE) + # don't want the legend
  theme_minimal() +
  theme(panel.background = element_rect(fill = "lightblue")) +
  coord_map(projection = "sinusoidal", # specify that we want to use the sinusoidal projection
            xlim = c(-180,180)) # set x limits
```

#### Making a map of California
```{r}
# use the states dataset
head(states)
```

Filter out just the California data
```{r}
CA_data<-states %>%
  filter(region == "california")

head(CA_data)
```

```{r}
ggplot() + 
  geom_polygon(data = CA_data, 
               aes(x = long, 
                   y = lat, 
                   group = group),
               color = "black", # add color to the lines
               fill = "turquoise") + 
  theme_void() +  # can remove all the axis (lat/long)
  coord_map() # default is Mercator
```

### Adding population density data!
```{r}
# Look at the county data we got earlier
head(counties)[1:3,] # only showing the first 3 rows for space
```

```{r}
# Look at county data
head(popdata)
```

```{r}
CApop_county<-popdata %>%
  select("subregion" = County, Population)  %>% # rename the county column
  inner_join(counties) %>% # join with counties
  filter(region == "california") # some counties have same names in other states so only keep california

head(CApop_county)
```

#### Map of CA population by county
```{r}
ggplot() + 
  geom_polygon(data = CApop_county,
               aes(x = long,
                   y = lat,
                   group=group,
                   fill = Population),
               color = "black") + 
  coord_map() +
  theme_void()
```

But LA has soo many people that it looks low pop everywhere else, so we make it log scale for easier interpretation
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10") # log transform data 
```

### Adding Stars Data
The stars dataset has the number of seastars per/m2 at different sites in California.

```{r}
head(stars)
```

```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  geom_point(data = stars, # add a point at all my sites of star counts
             aes(x = long,
                 y = lat,
                 size = star_no)) + # change the point size by number of stars
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10") +
  labs(size = "# stars/m2") 

ggsave(here("Week7","Output","Homework", "CApop.pdf"))
```

