---
title: "w7b_lecture"
author: "Micaela Chapuis"
date: "2024-10-13"
output: html_document
---

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(ggmap)
library(ggspatial) # to add scale bars and compass arrows
```

## Load Data
```{r}
ChemData <- read_csv(here("Week7", "Data", "chemicaldata_maunalua.csv"))

glimpse(ChemData)
```

## Get base maps from ggmap
get_map() is the function to get a basemap from Google maps.  

Location as a string
```{r}
Oahu <- get_map("Oahu")
```

Then use ggmap to plot the base layer
```{r}
ggmap(Oahu)
```

### Getting a specific place

We can also put in lat and lon coordinates to get a specific place as the **center of the map.**   
Get *Wailupe* coordinates
```{r}
# Make a data frame of lon and lat coordinates
WP <- data.frame(lon = -157.7621, lat = 21.27427) # coordinates for Wailupe

# Get base layer
Map1 <- get_map(WP)

# plot it
ggmap(Map1)
```

### Zooming in on a location
The *zoom* argument, an integer from **3 to 20** specifying how large the spatial extent should be around the center, with 3 being the *continent level* and 20 being roughly the *single building level*
```{r}
Map1 <- get_map(WP, zoom = 17)
ggmap(Map1)
```

### Change the Map Type
We can use Google Maps types and Stadia types
```{r}
Map1 <- get_map(WP, zoom = 17, maptype = "satellite")
ggmap(Map1)
```

Other option: (for some reason doesn't work but I don't actually like the stamen maps so it's fine)  
Needed to install Stadia API (https://docs.stadiamaps.com/tutorials/getting-started-in-r-with-ggmap/)
```{r}
#Map1<-get_map(WP,zoom = 17, maptype = "stamen_watercolor", source = "stadia")

#ggmap(Map1)
```


### Adding points and scale bar + north arrow
```{r}
Map1 <- get_map(WP, zoom = 17, maptype = "satellite")

ggmap(Map1) +
  geom_point(data = ChemData,
             aes(x = Long, y = Lat, color = Salinity),
             size = 4) +
  scale_color_viridis_c() +
  annotation_scale(bar_cols = c("black", "white"), # make colors black and white
                   location = "bl") + # put it in the bottom left
  annotation_north_arrow(location = "tl") + # add a north arrow in top left
  coord_sf(crs = 4326) # for the scale bar to work it needs to be in this coordinate system - this is a typical coordinate reference system for a GPS (WGS84)
```

## Geocoding
If we don't know the lat and long of the place we want, we can use geocode() to get exact locations that we can then use in maps
```{r}
geocode("the white house")
geocode("University of Hawaii at Manoa")
```

