---
title: "W15 Homework"
author: "Micaela Chapuis"
date: "2024-12-06"
output: github_document
---
```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


## Assignment Details
You have a set of 4 .csv files in data/homework. Each of these files is a timeseries of temperature and light data collected in tide pools in Oregon by Jenn Fields. Your goal is to bring in all 4 files and calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool. Use both a for loop and map() functions in your script. (Basically, do it twice).  

## Load Libraries
```{r}
library(here)
library(tidyverse)
```

## For Loop

First, let's read in a file to see what it looks like
```{r}
test <- read_csv(here("Week15", "Data", "homework","TP1.csv"))
glimpse(test)
```

Now make a list of all files in the specified path
```{r}
filepath <- here("Week15", "Data", "homework") # specify file path

files <- dir(path = filepath, pattern = ".csv") # make a list of all files in this path that match the pattern (in this case ".csv")

files # view
```


Pre-allocate space for the for loop: making an empty dataframe that has a row for each file (each file is a tidepool) and includes the file name (tidepool number), the mean and SD for temperature, and the mean and SD for light.
```{r}
pool_data <- tibble(filename =  rep(NA, length(files)),  # column for the file name (tidepool number), length is the number of files we're importing
                   mean_temp = rep(NA, length(files)), # column for the mean temperature
                   sd_temp = rep(NA, length(files)), # column for the temperature sd
                   mean_light = rep(NA, length(files)), # column for the mean light
                   sd_light = rep(NA, length(files))) # column for light sd
                   
pool_data
```

Now check that my code will work before putting it in the loop
```{r}
raw_data <- read_csv(paste0(filepath,"/",files[1])) # read in the first file to make sure it works
# paste0 gives filepath/filename.csv from our list of file names called 'files'
head(raw_data) # it works!
```

Also test our mean and sd code before putting it in loop
```{r}
mean_temp <- mean(raw_data$Temp.C, na.rm = TRUE) # calculate mean temp
mean_temp

sd_temp <- sd(raw_data$Temp.C, na.rm = TRUE) # calculate sd for temp
sd_temp
```

Actual for loop now!
```{r}
for (i in 1:length(files)){ # loop over the number of files listed in 'files'
  raw_data <- read_csv(paste0(filepath,"/",files[i])) # read in each file as it iterates over the number of files in the list (1-4)
  
  pool_data$filename[i] <- files[i] # store the name of each file in the filename colun
  
  
  pool_data$mean_temp[i] <- mean(raw_data$Temp.C, na.rm =TRUE) # calculate the mean temp for each file and store it in the corresponding row
  pool_data$sd_temp[i] <- sd(raw_data$Temp.C, na.rm = TRUE) # calculate temp sd for each file and store it in corresponding row
  
  
  pool_data$mean_light[i] <- mean(raw_data$Intensity.lux, na.rm = TRUE) # calculate mean light for each file and store it in corresponding row
  pool_data$sd_light[i] <- sd(raw_data$Intensity.lux, na.rm = TRUE) # calculate light sd for each file and store it in corresponding row
} 

pool_data
```

## Map Functions
Using the same filepath from above to store the full path of each file into 'files'
```{r}
files <- dir(path = filepath, pattern = ".csv", full.names = TRUE)
```

```{r}
pool_data_map <- files %>% # what we're iterating over, the full path to each file
                  set_names() %>% # sets the id of each row to the file name
                  map_df(read_csv, .id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  
                  group_by(filename) %>% # group by filename (aka by tidepool)
                  summarise(mean_temp = mean(Temp.C, na.rm = TRUE), # calculate mean temp
                            sd_temp = sd(Temp.C, na.rm = TRUE), # calculate temp sd
                            mean_light = mean(Intensity.lux, na.rm = TRUE), # calculate mean light
                            sd_light = sd(Intensity.lux, na.rm = TRUE)) # calculate light sd

pool_data_map
```

