---
title: "w7_homework"
author: "Micaela Chapuis"
date: "2024-10-09"
output: html_document
---

## Assignment Details
You need to make a map using RMarkdown with one of the following datasets from past Tidy Tuesdays.  
Make sure, as usual, that you scripts and output are saved appropriately.  

## Dataset Details
Using the Pet Cats UK dataset available [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md)

Original article:  
Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L, Tindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation. doi:10.1111/acv.12563

Movebank data package:  
McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of pet cats [United Kingdom]. Movebank Data Repository. doi:10.5441/001/1.pf315732

## Loading Libraries
```{r}
library(tidyverse)
library(here)
library(mapdata)
```


## Getting the Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2023-01-31')
tuesdata <- tidytuesdayR::tt_load(2023, week = 5)

cats_uk <- tuesdata$cats_uk
cats_uk_reference <- tuesdata$cats_uk_reference

# Or I could've read in the data manually:

# cats_uk <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk.csv')
# cats_uk_reference <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk_reference.csv')

```

### Looking at Data
```{r}
head(cats_uk)
head(cats_uk_reference)
```
First, I want to see how many different unique cats are included, so I use length to get the count of unique tag_ids. Because there are so many, I want to look at what their names are so I can then choose only some of them. 
```{r}
length(unique(cats_uk$tag_id)) # how many unique tags are there

unique(cats_uk$tag_id) # what are the tag/cats names
```

# Data Processing
Joining datasets, selecting columns I want to keep, and selecting specific cat
```{r}
cats <- left_join(cats_uk, cats_uk_reference, by = 'tag_id') %>%  # join datasets by tag id
        select(c("tag_id", "animal_id", "timestamp", "location_long", # select columns I want to keep
                 "location_lat", "animal_sex", "age_years")) %>% 
        filter(tag_id %in% c("Pants-Tag")) # selecting specific cats

#, "SmokeyLongnose-Tag", "Guinness-Tag", "Lightening Bugg-Tag"
```

Creating UK dataset from mapdata package
```{r}
uk <- map_data("world", region = "UK")
head(uk)
```

## Creating Map of the UK
```{r}
ggplot() + 
  geom_polygon(data = uk, 
               aes(x = long, 
                   y = lat, 
                   group = group),
               color = "black", # add color to the lines
               fill = "gray90") + 
  geom_point(data = cats, 
             aes(x = location_long,
                 y = location_lat)) +
  theme_minimal() +
  coord_map(xlim = c(-5.072, -5.069), ylim = c(50.1514, 50.154))
```

