---
title: "w7_homework"
author: "Micaela Chapuis"
date: "2024-10-09"
output: html_document
---

## Assignment Details
You need to make a map using RMarkdown with one of the following datasets from past Tidy Tuesdays.  
Make sure, as usual, that you scripts and output are saved appropriately. 

## Dataset Details
Using the Pet Cats UK dataset available [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md)

Original article:  
Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L, Tindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation. doi:10.1111/acv.12563

Movebank data package:  
McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of pet cats [United Kingdom]. Movebank Data Repository. doi:10.5441/001/1.pf315732


```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Loading Libraries
```{r}
library(tidyverse)
library(here)
library(ggmap)
library(ggspatial) # to add scale bars and compass arrows
library(gganimate) # to animate the points on the map
library(gifski)    # to render the animated map into a gif
```


## Getting the Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2023-01-31')
tuesdata <- tidytuesdayR::tt_load(2023, week = 5)

cats_uk <- tuesdata$cats_uk
cats_uk_reference <- tuesdata$cats_uk_reference

# Or I could've read in the data manually:

# cats_uk <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk.csv')
# cats_uk_reference <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk_reference.csv')

```

### Looking at Data
```{r}
head(cats_uk)
head(cats_uk_reference)
```
First, I want to see how many different unique cats are included, so I use length to get the count of unique tag_ids. Because there are so many, I want to look at what their names are so I can then choose only one of them (the one with the coolest name, obviously). 
```{r}
length(unique(cats_uk$tag_id)) # how many unique tags are there

unique(cats_uk$tag_id) # what are the tag/cats names
```

## Data Processing
### Joining datasets, selecting columns I want to keep, and selecting specific cat
```{r}
cats <- left_join(cats_uk, cats_uk_reference, by = 'tag_id') %>%  # join datasets by tag id
        select(c("tag_id", "animal_id", "timestamp", "location_long", "location_lat")) %>% # select columns I want to keep  
        filter(tag_id %in% "SmokeyLongnose-Tag") # selecting specific cat

# some favorites: "Pants-Tag", "SmokeyLongnose-Tag", "Guinness-Tag", "Lightening Bugg-Tag"
# fun fact: Smokey lives in a town called Penzance in the UK
```


### Getting Smokey's mean coordinates + Getting base layer centered around those 
```{r}
# Make a data frame of Smokey's mean lon and lat coordinates 
SmokeyMean <- data.frame(lon = mean(cats$location_long), lat = mean(cats$location_lat)) # mean coordinates

# Get base layer centered at those coordinates, zoom in, and make it satellite type map so we can see where Smokey goes
SmokeyMap <- get_map(SmokeyMean, zoom = 18, maptype = "satellite") # zoom 18 cuts out two points in the tree area at the bottom, but lets us see better how many houses Smokey goes to (probably to beg for food)
```


### Making the Map! 
```{r}
map <- ggmap(SmokeyMap) +  # plot map made above
       geom_point(data = cats, # add cat data
                  aes(x = location_long, 
                      y = location_lat,
                      group = seq_along(timestamp)), # this makes it so that the points already plotted stay on the map
                  color = "#EA4335", # using the color of the Google Maps marker
                  size = 3,
                  alpha = 0.7) + 
       theme_minimal() +
       labs(x = "Longitude", # set labels and title
            y = "Latitude",
            title = "Smokey Longnose's Journey") +
  
# Add scale bar
       annotation_scale(bar_cols = c("grey10", "white"), # add scale bar, make colors dark grey and white
                        location = "tr",  # put it in the top right
                        text_col = "white") + # make text color white so it can be seen
  
# Add North Arrow
       annotation_north_arrow(location = "tl",# add a north arrow in top left
                              height = unit(1, "cm"), # set size
                              width = unit(1, "cm"),
                              style = north_arrow_fancy_orienteering(text_col = 'white', # set type of arrow and colors for all its parts
                                                                     line_col = 'white',
                                                                     fill = 'white')) +
       coord_sf(crs = 4326) + # necessary crs for scale bar to work
  
# Animate plot by timestamp
       transition_reveal(timestamp) # this animates the plot over the days/times in the timestamp




# Render the animation
animate(map, #  set the duration, fps, width and height of the animated map
        duration = 15, 
        fps = 30, 
        width = 600, 
        height = 600, 
        renderer = gifski_renderer())

#Save the animated map into the Week 7 Output folder
anim_save(here("Week7","Output", "Homework", "SmokeyLongnosesJourney.gif"))
```
