---
title: "w7_homework"
author: "Micaela Chapuis"
date: "2024-10-09"
output: html_document
---

## Assignment Details
You need to make a map using RMarkdown with one of the following datasets from past Tidy Tuesdays.  
Make sure, as usual, that you scripts and output are saved appropriately.  

## Dataset Details
Using the Pet Cats UK dataset available [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md)

Original article:  
Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L, Tindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation. doi:10.1111/acv.12563

Movebank data package:  
McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of pet cats [United Kingdom]. Movebank Data Repository. doi:10.5441/001/1.pf315732


## Getting the Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2023-01-31')
tuesdata <- tidytuesdayR::tt_load(2023, week = 5)

cats_uk <- tuesdata$cats_uk
cats_uk_reference <- tuesdata$cats_uk_reference

# Or I could've read in the data manually:

# cats_uk <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk.csv')
# cats_uk_reference <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk_reference.csv')

```

### Looking at Data
```{r}
head(cats_uk)
head(cats_uk_reference)
```
First, I want to see how many different unique cats are included, so I use length to get the count of unique tag_ids
Because there are so many, I want to look at what their names are so I can then choose only some of them. 
```{r}
length(unique(cats_uk$tag_id)) # how many unique tags are there

unique(cats_uk$tag_id) # what are the tag/cats names
```

There are many columns I won't use so I only select the ones I want:
From cats_uk,  I want tag_id, timestamp, location_long, location_lat
From cats_uk_reference, I want tag_id (for joining the datasets), animal_id, and animal_sex, age_years
```{r}

```


Only select the cats I want (the ones with the best names, obviously)
"Pants-Tag", "SmokeyLongnose-Tag", "Guinness-Tag", "Lightening Bugg-Tag"