---
title: "week4b_lecture"
author: "Micaela Chapuis"
date: "2024-09-18"
output: html_document
---

# Week 4 Online Lecture

## Load libraries
```{r}
library(tidyverse)
library(here)
```

# Load data
```{r}
ChemData <- read_csv(here("Week4", "Data", "chemicaldata_maunalua.csv"))
View(ChemData)
```

Looking at it
```{r}
glimpse(ChemData)
```

```{r}
ChemData_clean <- ChemData %>% filter(complete.cases(.)) # filters out everything that isn't a complete row

ChemData_clean
```

### Separate Function
Tide_time has two types of data in it (low_day), but we want it in separate columns
```{r}
ChemData_clean <- ChemData %>%
  drop_na() %>% # filters out everything that isn't a complete row
  separate(col = Tide_time, # choosing the tide time col
           into = c("Tide", "Time"), # separate into two columns, tide and time
           sep = "_", # separate by the _
           remove = FALSE) # keeping the original tide_time col

ChemData_clean
```


### Unite
If we want to combine two columns into one
```{r}
ChemData_clean <- ChemData %>%
  drop_na() %>% # filters out everything that isn't a complete row
  separate(col = Tide_time, # choosing the tide time col
           into = c("Tide", "Time"), # separate into two columns, tide and time
           sep = "_", # separate by the _
           remove = FALSE) %>% # keeping the original tide_time col
  unite(col = "Site_Zone", # the name of the NEW col
          c(Site,Zone), # the columns to unite
        sep = ".", # let's put a . in the middle
        remove = FALSE) # keep the original

ChemData_clean
```


### Pivoting data between wide and long
Wide: one observation per row and all the different variables are columns
Long: one unique measurement per row and all the info about that measurement in the same row

wide to long: pivot_longer()
long to wide: pivot_wider()

#### Wide to Long
```{r}
ChemData_long <- ChemData_clean %>% 
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols
              names_to = "Variables",  # the names of the new cols with all the column names
              values_to = "Values") # names of the new column with all the values

ChemData_long
```

Now we can calculate the mean and variance for all variables at each site
```{r}
ChemData_long %>%
  group_by(Variables, Site) %>% # group by everything we want
  summarise(Param_means = mean(Values, na.rm = TRUE), # get mean
            Param_vars = var(Values, na.rm = TRUE)) # get variance
```

#### Challenge
Calculate mean, variance, and standard deviation for all variables by site, zone, and tide
```{r}
ChemData_long %>%
  group_by(Site, Zone, Tide) %>% # group by everything we want
  summarise(Param_means = mean(Values, na.rm = TRUE), # get mean
            Param_vars = var(Values, na.rm = TRUE), # get variance
            Param_sd = sd(Values, na.rm = TRUE)) # get sd
```

#### Facet with Long data
```{r}
ChemData_long %>% 
  ggplot(aes(x = Site, y = Values)) +
  geom_boxplot() +
  facet_wrap(~Variables, scales = "free") # w/o scales free all panels would have the same y axis scale, though they all have very different scales 
# free affects x and y. If we only want one we'd do free_x or free_y
```

#### Long to Wide
```{r}
ChemData_wide <- ChemData_long %>% 
  pivot_wider(names_from = Variables, #colums with the names for new column
              values_from = Values) # column with the values

ChemData_wide              
```

### Summary Statistics + Exporting
Start from the beginning and work through our entire flow again, ending with the data export

1) Remove all NAs
2) Separate Tide_time into two columns 
3) Pivot longer
4) Group by Variable, Site, Time, and calculate means
5) Convert back to wide
6) Export the csv file of your summary statistics using write_csv()
```{r}
ChemData_clean <- ChemData %>%
  drop_na() %>% # filters out everything that isn't a complete row
  separate(col = Tide_time, # choosing the tide time col
           into = c("Tide", "Time"), # separate into two columns, tide and time
           sep = "_", # separate by the _
           remove = FALSE) %>% # keeping the original tide_time col
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols
              names_to = "Variables",  # the names of the new cols with all the column names
              values_to = "Values") %>% # names of the new column with all the values
  group_by(Variables, Site, Time) %>% # variables by site and time
  summarise(mean_vals = mean(Values, na.rm = TRUE)) %>%
  pivot_wider(names_from = Variables,
              values_from = mean_vals) %>% #notice it is now mean_vals as the column name
  write_csv(here("Week4", "Output", "online_lecture_output.csv")) #export as csv to the right folder


ChemData_clean
```

