---
title: "W9a Homework"
author: "Micaela Chapuis"
date: "2024-10-23"
output: 
  github_document:
    toc: true
---

## Assignment Details

Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples per function showing its flexibility. 

```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
# set the path for figure output
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../../Output/Homework/")
```


## Load Libraries
```{r}
library(tidyverse)
library(here)
```

## Load Data
```{r}
intertidal <- read_csv(here("Week9", "Data", "intertidaldata.csv")) %>% glimpse()
```

## Data Cleaning
Boooo cleaning strings (part 1)  
Cleaning column names
```{r}
names(intertidal) <- names(intertidal) %>% # take column names
                     str_replace_all(pattern = "\\)|\\(", replacement = "") %>% # replace parentheses from col names with nothing
                     str_replace_all(" ", "_") # replace all spaces with _
```


Boooo cleaning strings (part 2)  
Cleaning Quadrat character data
```{r}
intertidal$Quadrat <- intertidal$Quadrat %>% 
                      str_replace_all(pattern = "\\.|[0-9]", replacement = "") %>% # replace periods or numbers with nothing
                      str_trim() # trim white spaces
```

For the plots, I need to order the Quadrat column Low, Mid, High
```{r}
intertidal$Quadrat <- factor(intertidal$Quadrat, levels=c("Low", "Mid", "High")) # make factor and put it into that specific order
```


## Function 1 
I want this function to calculate the average percent cover of different sessile inverts (and algae/rock) in the low, mid, or high intertidal zone at some specific site. (Yes, it can also do it for the mobile invert counts, but I couldn't think of another more inclusive name, so percent cover only it is.)
```{r}
mean_pcover <- function(data = intertidal, site, quadrat, variable) { # data is intertidal by default, then we give it a site, a quadrat (low, mid, high intertidal location) and which variable we want the mean of
  avg <- data %>% 
    filter(Site == site & Quadrat == quadrat) %>% # only select the specified site and quadrat
    summarise(average = mean({{variable}}, na.rm =TRUE)) # take the mean of the specified variable
  return(avg)
}
```

### Example 1: Mean percent cover of mussels in the mid intertidal at Scripps
```{r}
mean_pcover(site = "Scripps", quadrat = "Mid", variable = Mussels) # Scripps and Mid are characters because they are values in rows that we need to filter by, Mussels is a column
```

### Example 2: Mean percent cover of algae in the low intertidal at Bodega
```{r}
mean_pcover(site = "Bodega", quadrat = "Low", variable = Algae)
```

### Example 3: Mean percent cover of small barnacles in the high intertidal at Starfish Point
```{r}
mean_pcover(site = "Starfish Point", quadrat = "High", variable = Small_Barnacles)
```

## Function 2
Boxplot of percent cover of inverts at the different intertidal zones at a specific site
```{r}
pcover_plot <- function(data = intertidal, site, variable){ # data is intertidal by default, site and variable need to be specified
                data %>% 
                filter(Site == site) %>%  # filter by site
                ggplot(aes(x = Quadrat, y = {{variable}})) + # x axis is Quadrat column, y is the variable we specify
                  geom_boxplot() +
                theme_bw() + # theme
                labs(x = "Location on Intertidal")
}
```

### Example 1: Anemone percent cover at Scripps
```{r HW9a-plot1}
pcover_plot(site = "Scripps", variable = Anemone) + 
  labs(y = "Anemone Percent Cover")
```

### Example 2: Bare rock percent cover at Pyramid Point
```{r HW9a-plot2}
pcover_plot(site = "Pyramid Point", variable = Bare_Rock) + 
  labs(y = "Bare Rock Percent Cover")
```

### Example 3: Gooseneck barnacles percent cover at Cape Arago
```{r HW9a-plot3}
pcover_plot(site = "Cape Arago", variable = Gooseneck_Barnacles) + 
  labs(y = "Gooseneck Barnacles Percent Cover")
```


