---
title: "W9a Homework"
author: "Micaela Chapuis"
date: "2024-10-23"
output: 
  github_document:
    toc: true
---

## Assignment Details

Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples per function showing its flexibility. Due, as always, at 1pm on Tuesday.

```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
# set the path for figure output
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


## Load Libraries
```{r}
library(tidyverse)
library(here)
```

## Load Data
```{r}
intertidal <- read_csv(here("Week9", "Data", "intertidaldata.csv")) %>% glimpse
```

I want to see the average cover/counts of some variable in the low, mid, or high intertidal zone at some specific site.
```{r}
mean_pcover <- function(data, site, quadrat, variable){
  #average <- data %>% 
   #          filter(Site == site && Transect == transect) %>%
  avg <- data %>% 
    filter(Site == site & Quadrat == quadrat) %>%
    summarise(average = mean({{variable}}, na.rm =TRUE))
  return(avg)
}
```

```{r}
mean_pcover(intertidal, "Scripps", "Mid", Mussels)
```

