---
title: "W9b Homework"
author: "Micaela Chapuis"
date: "2024-10-23"
output: 
  github_document:
    toc: true
---

## Assignment Details

Working with intertidal data. Today you will need to use skills that you have learned in working with words to clean up some issues with the character data and use what you learned today to reorder your factors into orders that make sense. (Also the column headers could be better...). HINT: New package idea to fix this is using the janitor package.  

Your goal, as per usual, is to make a plot, any plot where one of the axes is a factor. Save everything in the appropriate folders.  

You have two possible files to play with in the data folder (you don't HAVE to use both): intertidal.csv and intertidal_latitude.csv (as well as a data dictionary).  

Importantly, if you decide to visualize anything by tide height it should go from low to mid to high tide on the plot and if you decide to visualize by site it should go in order of latitude (e.g. south to north or north to south).  

```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load Libraries
```{r}
library(here)
library(tidyverse)
library(janitor)
library(ggspatial)
```

## Load in Data
```{r}
intertidal_data <- read_csv(here("Week9", "Data", "intertidaldata.csv")) %>% glimpse()
intertidal_sites <- read_csv(here("Week9", "Data", "intertidaldata_latitude.csv")) %>% glimpse()
```

## Joining data by site name
```{r}
intertidal <- left_join(intertidal_data, intertidal_sites)
```


## Cleaning Data
Oops I did this in the previous assignment, but I get to use the janitor package this time!  

### Cleaning up the Quadrat strings (Low, Mid, High)
```{r}
intertidal$Quadrat <- intertidal$Quadrat %>% 
                      str_replace_all(pattern = "\\.|[0-9]", replacement = "") %>% # replace periods or numbers with nothing
                      str_trim() %>% # trim white spaces
                      factor(levels=c("Low", "Mid", "High")) # make factor and put levels in specific order
```

### Cleaning up the Column Names
```{r}
intertidal <- clean_names(intertidal) # makes everything separated by _ and lowercase

# well that was easier, should've done this homework first
```


## Plot
Boxplot of algae percent cover in the mid intertidal by site (ordered by latitude) with a north arrow
```{r}
intertidal %>% 
  filter(quadrat == "Mid") %>% # only use Mid intertidal data
  ggplot(aes(x = algae, 
             y = fct_reorder(site, latitude))) + # order the sites by latitude (ascending)
    geom_boxplot() +
    theme_bw() +
    labs(x = "Algae Percent Cover",
         y = "Site") +
    annotation_north_arrow(location = "tr", # put a north arrow top right
                           height = unit(1, "cm"), # set arrow size
                           width = unit(1, "cm")) +
    theme(axis.title = element_text(size = 12), # making all the text bigger
          axis.text = element_text(size = 10))


ggsave(here("Week9", "Output", "Homework", "w9b_homework.jpeg"))
```

