---
title: "W14 Tidy Tuesday"
author: "Micaela Chapuis"
date: "2024-11-27"
output: github_document
---

```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(cowsay)
```

## Load Data
```{r}
cbp_resp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-26/cbp_resp.csv')
```

## Data Wrangling
First, I want to see which of areas of responsibility have fewer countries of citizenship observed so I'm not working with such a huge dataset (you'll thank me later)
```{r}
cbp_resp %>% 
  group_by(area_of_responsibility) %>% # for each area of resp
  summarize(count = length(unique(citizenship))) %>%  # how many unique countries were observed
  view() # open it because it's long
```

Now, focusing on our specific area of responsibility (Spokane Sector, because it has the fewest countries observed), I want to see how many individuals encountered are from each country
```{r}
data <- cbp_resp %>% 
        filter(area_of_responsibility == "Spokane Sector") %>% # filter to only see encounters in Spokane
        select(citizenship, encounter_count) %>% # keep only the columns I care about
        group_by(citizenship) %>% # group by country of citizenship
        summarize(encounter_count = sum(encounter_count)) # add up number of individuals encountered from each country
```

## Output!
My new thing this week is using the 'cowsay' package (and the sample function too I guess). I will have a randomly chosen animal say "country of citizenship: # individuals" for each of the countries of citizenship identified (only 13 countries in Spokane so it doesn't get too crazy)

```{r}
for(i in 1:nrow(data)){  # for each row of the df data
  animal_chosen <- sample(names(animals), 1) # pick a random animal name from the list of possible cowsay animals and store it into "animal_chosen"
  say(what = paste0(data$citizenship[i], ": ", data$encounter_count[i]), # say the country of citizenship at that row, then ": " and then the number of encounters at that row
      by = animal_chosen) # the randomly chosen animal who gets to say that
}
```

