---
title: "w5_homework"
author: "Micaela Chapuis"
date: "2024-09-29"
output: html_document
---

# Week 5 Homework

* Read in conductivity and depth data
* Convert date columns appropriately
* Round the conductivity data to the nearest 10 seconds so it matches with the depth data
* Join the two dataframes together (with no NAs, so a join where only exact matches between the two dfs are kept)
* Take averages of date, depth, temp, and salinity by minute
* Make any plot using the averaged data

## Loading libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
```

## Loading in data
```{r}
CondData <-read_csv(here("Week5","Data", "CondData.csv"))
DepthData <- read_csv(here("Week5", "Data", "DepthData.csv"))
```

Look at data
```{r}
CondData # date is a character
DepthData # date is a datetime already
```


## Converting date columns + rounding CondData 
Well, just one column, the other one is already a datetime
```{r}
CondData <- CondData %>% 
            mutate(date = mdy_hms(date)) %>%  # converts to datetime 
            mutate(date = round_date(date, unit = "10 second")) #rounds to the nearest 10 seconds
  
#DepthData <- DepthData %>%       # it's already a datetime but this is what code would be if I needed it
#             mutate(date = mdy_hms(date)) 

CondData # view
DepthData # view
```

## Join Dataframes
Using inner join because it keeps only complete data in both datasets
```{r}
all_data <- inner_join(CondData, DepthData) # joins by date because it's the only thing they have in common
# only keeping the rows where all things were recorded at the same times

all_data # view
```


## Taking averages of the variables by minute + plotting
```{r}
all_data %>% select(date, Temperature, Salinity, Depth) %>% # we don't need Serial or Absolute Pressure
  
             mutate(minutes = minute(date)) %>% # make a new column with minutes extracted
  
  # Pivoting to long to make taking average easier
             pivot_longer(cols=c(Temperature:Depth), # columns I want to pivot
                          names_to = "Parameter", # name of the new column with variable names
                          values_to = "Value") %>% # name of new column with values
  
             group_by(minutes, Parameter) %>% # group by each unique minute and each parameter
 
             summarise(Mean = mean(Value)) %>% # take the mean by minute of each parameter 
  
  # Make Plot
             ggplot(aes(x= minutes, y = Mean, color = Parameter)) +
                geom_point(size = 1, alpha = 0.8) + # using points
                geom_line() +  # connect points with lines
                facet_wrap(~Parameter, 
                           ncol =1, # horizontally faceted by Parameter
                           scales = "free", # they each have their own y scale
                           labeller = labeller(Parameter = c(`Depth` = "Depth (m)", # set facet title names (assuming units)
                                                             `Salinity` = "Salinity (ppt)",
                                                             `Temperature` = "Temperature (°C)"))) + 
                scale_color_manual(values = c("#125B9A", "#86D293", "#F05A7E")) +  # set colors manually
                theme_bw() + # theme
                labs(x = "Minute",   # set labels
                     y = "Average Value (Note different scales)",
                     title = "Depth, Salinity and Temperature Averaged by Minute") +
                theme(axis.title = element_text(size = 12), # making all the text bigger
                      axis.text = element_text(size = 11),
                      legend.position = "none") + # don't need this legend because we have the facet titles
                scale_x_continuous(breaks = seq(from = 0, to = 60, by = 10)) # setting the x axis ticks every 10 mins


ggsave(here("Week5", "Output", "Homework", "w5_homework.jpeg"), width = 7, height = 5)
```

