---
title: "w13_lecture"
author: "Micaela Chapuis"
date: "2024-11-19"
output: github_document
---
```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(reprex) # reproducible example
library(datapasta) # copy and paste
library(styler) # copy and paste in style
```

By running code with {reprex}, you can create a reproducible example to post to Stack Overflow, Slack, GitHub, or an email to your collaborator.

## Making a reprex
Let's say we are trying to make a plot with the mpg dataset and you hit an error.
```{r}
mpg %>%
  ggplot(aes(x = displ, y = hwy))%>%
  geom_point(aes(color = class))
```

## Using real data
Open the stars.csv file and copy and paste the first few lines into R.

lat,long,star_no
33.548,-117.805,10
35.534,-121.083,1
39.503,-123.743,25
32.863,-117.24,22
33.46,-117.671,8

This is not helpful and won't run...

{datapasta} helps you copy and past snippets of your data in a reproducible way.
This time copy the code and got to Addins -> paste as tribble and then name it something.

tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L
  )
```{r}
data <- tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L
  )
```

You can now use this with {reprex} to make a reproducible example that can fully run with your own data and code.

----

```{r}
stars <- read_csv(here("Week13", "Data", "stars.csv"))
```



```{r}
library(tidyverse)

stars <- tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L
  )

ggplot(stars, aes(x = latitude, y = star_no)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of stars across latitude",
       x = "Latitude",
       y = "Number of Stars")

```



## Fixing classmates' code
```{r}
head(diamonds)
#> Error in head(diamonds): object 'diamonds' not found

diamonds%>% # don't need to both pipe the data and include it in the ggplot()
ggplot(aes(x = carat, y = price)) + 
      geom_point(aes(color = clarity), alpha = 0.5) +  # color by clarity
      labs(title = paste("Price vs Carat for"), #needed chatgpt, somehow this makes the reactive labels 
           x = "Carat", 
           y = "Price") +
      theme_minimal()
#> Error in diamonds %>% ggplot(diamonds, aes(x = carat, y = price)): could not find function "%>%"
```
