---
title: "w5a_lecture"
author: "Micaela Chapuis"
date: "2024-09-24"
output: html_document
---
# Week 5 - Tuesday Lecture
Today we are going to practice joins with data from Becker and Silbiger (2020).

## Load libraries
```{r}
library(tidyverse)
library(here)
```

## Load Data
```{r}
# Environmental data from each site
EnviroData<-read_csv(here("Week5","Data", "site.characteristics.data.csv"))
#Thermal performance data
TPCData<-read_csv(here("Week5","Data","Topt_data.csv"))
```

## Look at the data
```{r}
glimpse(EnviroData)
```


```{r}
glimpse(TPCData)
```
Site Letter is what links the two datasets together


## Pivot the Data
```{r}
EnviroData_wide <- EnviroData %>%
  pivot_wider(names_from = parameter.measured, # pivot the data wider
              values_from = values) %>%
  arrange(site.letter) # sort by alphabetical order for site letter
# otherwise was ordered by Site Name alphabetical order

EnviroData_wide
```

## Left Join
Joining both datasets by Site letter
```{r}
FullData_left <- left_join(TPCData, EnviroData_wide) # automatically joins by site.letter, but if labeled differently we would need to specify

FullData_left
```

Relocating so that all character (site info, coral ID) variables are first, and then all the parameters and numeric stuff
```{r}
FullData_left<- FullData_left %>%
   relocate(where(is.numeric), .after = where(is.character)) # relocate all the numeric data after the character data

FullData_left
```


Calculate means and variances of all collected (TPC and environmental) data by site

First pivot to long
Then group by site and variable, and use summarise to get the means and variances of the values for each variable

```{r}
all_means_vars <- FullData_left %>% 
  pivot_longer(cols = c(E:Topt, light:substrate.cover), # the cols you want to pivot. This says select every column from E to Topt and from light to susbtrate cover (because site.block was in between them)
              names_to = "Parameter",  # the names of the new cols with all the column names
              values_to = "Values") %>% # names of the new column with all the values
  group_by(site.letter, Parameter) %>%  # group by site and variables
  summarise(Mean = mean(Values), # get the means of all the values
            Variance = var(Values)) # get the variance of all the values

all_means_vars
```


Different way to do it
```{r}
FullData_summary <- FullData_left %>%
  group_by(site.letter) %>%
  summarise_at(vars(E:Topt, light:substrate.cover), 
               funs(mean=mean, var=var), na.rm = TRUE)

FullData_summary
```

Third option
```{r}
option3 <- FullData_left %>%
  group_by(site.letter) %>%
  summarise(across(where(is.numeric),  # summarises everything that is numeric
                   list(mean = mean, var = var), na.rm = TRUE)) # created a list, named columns mean and var

option3
```

Fourth Option
```{r}
option4 <- FullData_left %>%
  select(-site.block) %>% # here we don't specify which columns we want, so we need to specify if we want to remove a column that is numeric but shouldn't be summarised
  group_by(site.letter) %>%
  summarise_if(is.numeric,
               funs(mean=mean, var=var),
               na.rm = TRUE)

option4
```

## Creating a tibble
```{r}
# Make 1 tibble
T1 <- tibble(Site.ID = c("A", "B", "C", "D"), 
             Temperature = c(14.1, 16.7, 15.3, 12.8))
T1
```

```{r}
T2 <-tibble(Site.ID = c("A", "B", "D", "E"), 
            pH = c(7.3, 7.8, 8.1, 7.9))
T2
```

## Left vs Right Join
```{r}
left_join(T1, T2) # loosing E because it's only in T2
```


```{r}
right_join(T1, T2) # loosing C because it's only in T1
```

## Inner Join vs Full Join
```{r}
inner_join(T1, T2) # keeps only complete data in both datasets
```


```{r}
full_join(T1, T2) # keeps everything
```


## Semi Join vs Anti Join
```{r}
semi_join(T1, T2)  #keeps all rows from the first data set where there are matching values in the second data set, keeping just columns from the first data set.
```

```{r}
anti_join(T1, T2) # keeps all rows in the first data set that do not match anything in the second data set. This can help you find possible missing data across datasets.
# shows us that temperature is the only thing missing from site C

anti_join(T2, T1) # if flipped, results are different
# shows us that pH is the only thing missing from E
```

