---
title: "w5b_lecture"
author: "Micaela Chapuis"
date: "2024-09-29"
output: html_document
---

# Week 5 - Online Lecture
Using lubridate, converting and manipulating dates and times

Loading libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
```

## Some functions
```{r}
now() # current date and time
now(tzone = "EST") # can specify timezones
now(tzone = "GMT")
```

```{r}
today() # today's date 
today(tzone = "GMT") # can also specify timezones
```

```{r}
am(now()) # is it morning?
leap_year(now()) # is it a leap year?
```
 

## Date specifications for lubridate

- Dates must be a character
- Will get error if we try to convert a date when it's a factor

### Converting characters into ISO date format
```{r}
ymd("2021-02-24") # converts character into ymd date
```

```{r}
mdy("February 24 2021") # converts character into mdy date
```

```{r}
dmy ("24/02/2021") # converts character into dmy date
```

### Date and Times
```{r}
ymd_hms("2021-02-24 10:22:20 PM")
mdy_hms("02/24/2021 22:22:20")
mdy_hm("February 24 2021 10:22 PM") # didn't give seconds so output is 0 seconds
```


## Extracting specific date or time elements from datetimes
```{r}
# make a character string
datetimes <- c("02/24/2021 22:22:20",
               "02/25/2021 11:21:10",
               "02/26/2021 8:01:52")
```

```{r}
#convert to datetimes
datetimes <- mdy_hms(datetimes)
```

```{r}
# extract stuff
month(datetimes, label = TRUE) # label makes it so it gives us month name rather than number (makes month a factor!)

month(datetimes, label = TRUE, abbr = FALSE) # will give full month name rather than abbreviation 
```

```{r}
day(datetimes) # extract day
wday(datetimes, label = TRUE) # extract day of the week (with name)
```

```{r}
hour(datetimes)
minute(datetimes)
seconds(datetimes)
```


## Adding dates and times
```{r}
datetimes + hours(4) # adds 4 hours 
# notice it's hourS, not hour as before
```
hour() extracts the hour component from a time and hours() is used to add hours to a datetime

```{r}
datetimes + days(2) # adds 2 days 
# same as hours, days() != day()
```

Can do this ^ with minutes(), seconds(), months(), years(), etc


## Rounding dates
```{r}
round_date(datetimes, "minute") # round to nearest minute
round_date(datetimes, "5 mins") # round to nearest 5 minute
```

## Challenge
Read in conductivity data (CondData.csv) and convert the date column to a datetime. 
Data were collected every 10 seconds. 

### load in data
```{r}
CondData <-read_csv(here("Week5","Data", "CondData.csv"))

CondData # view
# Date is already a character
```

```{r}
CondData <- CondData %>% 
            mutate(date = mdy_hms(date))
CondData
```

