---
title: "week4b_homework"
author: "Micaela Chapuis"
date: "2024-09-18"
output: html_document
---

## Code Description
This script contains the code for Assignment 4b.

The assignment details are below: 
> Using the chemistry data:
 *Create a new clean script
 *Remove all the NAs
 *Separate the Tide_time column into appropriate columns for analysis
 *Filter out a subset of data (your choice)
 *Use either pivot_longer or pivot_wider at least once
 *Calculate some summary statistics (can be anything) and export the csv file into the output folder
 *Make any kind of plot (it cannot be a boxplot) and export it into the output folder
 *Make sure you comment your code and your data, outputs, and script are in the appropriate folders


## Loading libraries
```{r}
library(tidyverse)
library(here)
library(GGally) # for our correlogram
```

# Load data
```{r}
ChemData <- read_csv(here("Week4", "Data", "chemicaldata_maunalua.csv"))
```

Looking at it for inspiration
```{r}
glimpse(ChemData)
```

## Data Cleaning
I personally hate really long chunks with a lot of functions going on so I will be breaking stuff up into separate chunks 


Removing NAs and separating Tide_time column into two (Tide and Time)
```{r}
ChemData_clean <- ChemData %>%
  drop_na() %>% # filters out all rows that aren't complete (so filters out NAs)
  separate(col = Tide_time, # choosing the tide_time column to separate
           into = c("Tide", "Time"), # say what two columns we want to separate it into (tide and time)
           sep = "_", # use the _ as the separator
           remove = FALSE) # this makes it so we keep the original tide_time column just in case

ChemData_clean #viewing it just to check it
```

Filtering out a subset of the data: keeping only the Day data
```{r}
ChemData_day <- ChemData_clean %>% filter(Time == "Day") #keeping only rows of data collected during the Day
ChemData_day # viewing
```

## Analysis

I only want Temp, pH, and TA so I will select first just to make the output not too messy - storing it into its own df because I will use it later
```{r}
ChemData_day_select <- ChemData_day %>%
  select(Site, Season, Tide, Time, Temp_in, pH, TA)# keeping only the variables I want plus things I want to group by later (site, season, tide) and keeping time just for fun even though it's only Day

ChemData_day_select
```

Data is currently in Wide format, I will pivot to long for summary statistics
```{r}
ChemData_day_long <- ChemData_day_select %>%
  pivot_longer(cols = c(Temp_in, pH, TA), # the columns I want to pivot (temp, pH, TA)
              names_to = "Variables",  # the names of the new column with the variable (previous columns) names
              values_to = "Values") # names of the new column with all the values

ChemData_day_long
```

Summary Statistics: 
I want the average of temperature, pH and TA by tide, season, and site (and only for Day times but we already filtered that earlier)
For each Site at each season at each tide (low vs high) combination we'll get 3 rows (mean, var and sd of temp, pH, and TA)
We will export this as a csv into the output folder
```{r}
ChemData_day_long %>%
  group_by(Site, Season, Tide, Variables) %>% # group by everything we want (Site, Season, Tide)
  summarise(mean_vals = mean(Values, na.rm = TRUE), # get mean
            var_vals = var(Values, na.rm = TRUE), # get variance
            sd_vals = sd(Values, na.rm = TRUE)) %>% # get sd
  write_csv(here("Week4", "Output", "w4b_homework_summary_output.csv")) #export as csv to the right folder
```


## Plot! 
We are using Day only data, and we only want to look at relationships between Temperature, pH and TA (by site)
Using package "GGally"
Following: https://unc-libraries-data.github.io/R-Open-Labs/Extras/ggally/ggally.html

```{r}
ChemData_day_select %>%
  ggpairs(columns = c("Temp_in", "pH", "TA"),  # this specifies which variables we want in our columns
          ggplot2::aes(color = Site), # color by site
          diag = list(continuous = wrap("densityDiag", alpha = 0.6)), # diagonal is density plot, with some transparency bc of the overlap
          upper = list(continuous = "blank"), # the plots above the diagonal will just be blank
          lower = list(continuous = wrap("smooth", method = "lm", alpha = 0.6)), # plots below diagonals will be a scatterplot with a geom_smooth by Site
          columnLabels = c("Temperature", "pH", "Total Alkalinity"), # change names of columns
          legend = 1, # this makes the legend appear I guess? 
          switch = "both") + # switches labels from top/right to bottom/left
  theme_bw() + 
  labs(title = "Relationships between Temperature, pH, and Total Alkalinity in the Daytime" ) +
  scale_color_manual(values = c("#2a2550", "#e04d01")) + # changes color of the points in scatterplots
  scale_fill_manual(labels = c("Black Point", "Wailupe"), values = c("#2a2550", "#e04d01"))  # set labels for legend, changes color of the density plots

ggsave(here("Week4", "Output", "w4b_homework_plot.jpeg"), width = 7, height = 5)
```

