---
title: "W15 Lecture"
author: "Micaela Chapuis"
date: "2024-12-03"
output: github_document
---

## Load Libraries
```{r}
library(tidyverse)
library(here)
```

## For Loops!

### Simple for loops
Start with simple code  
```{r}
print(paste("The year is", 2000))
```

Now put it in a for loop  
```{r}
years <- c(2015:2021)

for (i in years){ # set up the for loop where i is the index
  print(paste("The year is", i)) # loop over i
}
```

What we just did printed something over and over, but it did not save it anywhere. Let's say you want to save a new vector with all the years. To do this we need to pre-allocate space and tell R where it is going to be saved.   

Create an empty dataframe called year_data with columns for year and year_name .  
```{r}
#Pre-allocate space for the for loop in an empty matrix that is as long as the years vector
year_data <- tibble(year =  rep(NA, length(years)),  # column name for year, repeating NA for the length of the sequence
                    year_name = rep(NA, length(years))) # column name for the year name
year_data
```

Now add the for loop  

Let's first add in the column that is going to have all the names in it. Notice that I added an index i in the column name.   
I also am having the index go from 1:length(years), which is 1:7. I use length() because it allows me to change the number of years in the vector without having to change the for loop.  
Then fill in the year column too.   
```{r}
for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i] <- paste("The year is", years[i]) # loop over i
  year_data$year[i]<-years[i] # loop over year
}
year_data
```

###  Using loops to read in multiple .csv files  
Let's say you have multiple data files where you want to perform the same action to each. You can use a for loop to do this.  

From the cond_data folder: using 3 files of salinity and temperature data collected from Mo'orea from a spatial survey.  

Read in one of the files so that you can see what it looks like.  
```{r}
testdata <- read_csv(here("Week15", "Data", "cond_data","011521_CT316_1pcal.csv"))
glimpse(testdata)
```

List files in a directory
```{r}
# point to the location on the computer of the folder
CondPath <- here("Week15", "Data", "cond_data")

# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename
# we could use regex to be more specific if you are looking for certain patterns in filenames
files <- dir(path = CondPath, pattern = ".csv")
files
```


Pre-allocate space for the loop  
Let's calculate the mean temperature and salinity from each file and save it  
```{r}
# make an empty dataframe that has one row for each file and 3 columns
 cond_data <- tibble(filename =  rep(NA, length(files)),  # column name for the file name
                   mean_temp = rep(NA, length(files)), # column name for the mean temperature
                   mean_sal = rep(NA, length(files))) # column name for the mean salinity
                   
cond_data
```

For loop time!  

First make sure code works before putting it in the loop
```{r}
raw_data <- read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works
# paste0 means pasting with no spaces
# / because it's path then backslash then file name
head(raw_data)
```
And write code to calculate the mean and test it first
```{r}
mean_temp <- mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean
mean_temp
```


Turning it into a for loop  

First just adding the data in
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files
  raw_data <- read_csv(paste0(CondPath,"/",files[i])) # read in all files, one at a time
  glimpse(raw_data)
}
```

Now add in the file name to the appropriate column
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
  raw_data <- read_csv(paste0(CondPath,"/",files[i]))
  #glimpse(raw_data)
  cond_data$filename[i] <- files[i]
} 

cond_data
```

Now add in our means calculations
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
  raw_data <- read_csv(paste0(CondPath,"/",files[i]))
  #glimpse(raw_data)
  
  cond_data$filename[i] <- files[i]
  
  cond_data$mean_temp[i] <- mean(raw_data$Temperature, na.rm =TRUE)
  cond_data$mean_sal[i] <- mean(raw_data$Salinity, na.rm =TRUE)
} 
cond_data
```

## Purrr Package  
**Map Functions**  
The pattern of looping over a vector, doing something to each element and saving the results is so common that the purrr package provides a family of functions to do it for you. There is one function for each type of output:  
  
- map() makes a list.   
- map_lgl() makes a logical vector.  
- map_int() makes an integer vector.  
- map_dbl() makes a double vector.  
- map_chr() makes a character vector.  
- map_df() makes a dataframe  
  
Each function takes a vector as input, applies a function to each piece, and then returns a new vector that’s the same length (and has the same names) as the input.  

Don't have to pre-allocate space!  

### Simple examples
There 3 ways to do the same thing in a map() function:  

- Use a canned function that already exists  

Let's calculate the mean from a set of random numbers and do it 10 times  

Create a vector from 1:10
```{r}
1:10 # a vector from 1 to 10 (we are going to do this 10 times)
```

for each time 1:10 make a vector of 15 random numbers based on a normal distribution
```{r}
1:10 %>% # the vector to iterate over
  map(rnorm, n = 15) # calculate 15 random numbers based on a normal distribution in a list
```
Calculate the mean from each list
```{r}
1:10 %>% # the vector to iterate over
  map(rnorm, n = 15)  %>% # calculate 15 random numbers based on a normal distribution in a list 
  map_dbl(mean) # calculate the mean (canned function that already exists). It is now a vector which is type "double"
```


### Same thing different notation
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function
  map_dbl(mean)
```

Use a formula when you want to change the arguments within the function
```{r}
1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  # using the ~ and the .x is used to represent the 'i'
  map_dbl(mean)
```

### Bring in files using purrr instead of a for loop

First find the files
```{r}
# point to the location on the computer of the folder
CondPath <- here("Week15", "Data", "cond_data")
files <- dir(path = CondPath,pattern = ".csv")
files
```

Or, we can get the full file names in one less step by doing this...
```{r}
files <- dir(path = CondPath, pattern = ".csv", full.names = TRUE)
# full names = TRUE gives the entire path for each one, so save the entire path name
files
```

Next, read in the files using map instead of a for loop while retaining the filename as a column.
```{r}
data <- files %>% # what we're iterating over, the full path to each file
          set_names()%>% # sets the id of each list to the file name
          map_df(read_csv, .id = "filename") # map everything to a dataframe and put the id in a column called filename
data
```

Now we have a regular dataframe and we can calculate means in the way we already know how! group_by filename and use summarize
```{r}
data <- files %>%
          set_names()%>% # set's the id of each list to the file name
          map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
          group_by(filename) %>% # pipe into whatever other code we need to use
          summarise(mean_temp = mean(Temperature, na.rm = TRUE),
                    mean_sal = mean(Salinity,na.rm = TRUE))
data
```

