---
title: "Week 6 Homework"
author: "Micaela Chapuis"  
output: rmdformats::downcute
---

## Assignment Details

-   Take any data and create at least one table and one figure
-   Create a beautiful RMarkdown output → **Knit the file!**
-   Have headings with clear explanations
-   Use bold, italics, and lists at least once

```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
# set the path for figure output
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../../Output/Homework/")
```

## Load Libraries

```{r}
library(palmerpenguins)
library(tidyverse)
library(reactablefmtr)
```

## Looking at data

```{r}
glimpse(penguins)
```

## Preparing Data for Table and Figure

Taking means of each variable by species and storing it in *penguin_species_means* because I **need** it for both my table and my figure

```{r}
penguin_species_means <- penguins %>% 
  # pivot to long to take the mean of each variable easily
  pivot_longer(cols=c(bill_length_mm:body_mass_g), # columns I want to pivot
                          names_to = "Variable", # name of the new column with variable names
                          values_to = "Value") %>%
  
  # group by species and take mean of each variable
  group_by(species, Variable) %>%
  summarise(Mean = mean(Value, na.rm = TRUE))
```

## Creating Table

Making a table of penguin data where we can *order the data in ascending or descending order*.\
The data is the average of each variable recorded (bill depth and length, body mass, and flipper length) by species (prepared above).\
The data will be colored by value, with higher values being a darker pink and lower values getting lighter.

Details:

-   Using the **{reactablefmtr}** package and the "nytimes" theme.

-   Details about the package are [here.](https://kcuilla.github.io/reactablefmtr/index.html)

-   The examples I followed are [here.](https://kcuilla.github.io/reactablefmtr/articles/themes.html)

```{r}
penguin_species_means %>%
  # pivot back to wide because I need the variables to be columns for my table
  pivot_wider(names_from = Variable, #colums with the names for new column
              values_from = Mean) %>%
  
  ## Creating table
  reactable(
    theme = nytimes(header_font_color = "black", # set theme for the table and specify color for header (column names)
                    font_color = "grey30"), # set font color to be a bit lighter
    
    # add column group header (Variable) above the variables
    columnGroup = list(
      colGroup(name = "Variable", 
               columns = c("bill_depth_mm","bill_length_mm", "body_mass_g", "flipper_length_mm"))),
    
    # for each column: set the name and align it, set column width, round the number, and set colors based on numbers
    columns = list(
      bill_depth_mm = colDef(name = "Bill Depth (mm)", # set the name of the column
                             maxWidth = 100, # set column width
                             align = "center", # center column name text
                             cell = function(x) # this is to specify I'm manipulating the cells
                                sprintf("%0.1f", x), # formats the number (rounded to 1 decimal)
                             style = color_scales(penguins, colors = c("white", "#fd84a9"))), # set color scale (pink for higher values, get closer to white with lower values)
      bill_length_mm = colDef(name = "Bill Length (mm)", # same as above
                              maxWidth = 100,
                              align = "center",
                              cell = function(x)
                                sprintf("%0.1f", x),
                              style = color_scales(penguins, colors = c("white", "#fd84a9"))),
      body_mass_g = colDef(name = "Body Mass (g)", # same as above
                           maxWidth = 100,
                           align = "center",
                           cell = function(x)
                               sprintf("%0.1f", x),
                           style = color_scales(penguins, colors = c("white", "#fd84a9"))),
      flipper_length_mm = colDef(name = "Flipper Length (mm)", # same as above
                                 maxWidth = 100,
                                 align = "center",
                                 cell = function(x)
                                    sprintf("%0.1f", x),
                                 style = color_scales(penguins, colors = c("white", "#fd84a9"))),
      species = colDef(maxWidth = 250, # make this column wider
                       align = "center", # center text
                       footer = "Source: palmerpenguins package", # add a footer with data source
                       footerStyle = list(color = "#999999"))) # make footer text a lighter color
  )
```

## Making a Figure

Making a *Lollipop Chart* ([example here](https://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html))

```{r w6_homework}
penguin_species_means %>%
  ggplot(aes(x=species, y=Mean, color = species)) + 
  facet_wrap(~Variable, # facet by variable
             scales = "free", # let each variable have its own scale
             labeller = labeller(Variable = c(`bill_depth_mm` = "Bill Depth (mm)", # set facet names
                                              `bill_length_mm` = "Bill Length (mm)",
                                              `body_mass_g` = "Body Mass (g)",
                                              flipper_length_mm = "Flipper Length (mm)"))) +
  geom_point(size=5) + 
  geom_segment(aes(x=species, # geom segment draws lines between points (x,y) and (x end, y end)
                   xend=species,  # x and x end are both just species because we don't want the line to move horizontally
                   y=0, 
                   yend=Mean), # vertically, line goes from the 0 until the value
               size = 1.5) + # making lines thicker
  labs(x = "Species", # axis labels
       y  = "Mean (Scales vary)") +
  theme_bw() +
  theme(legend.position = "none", # removing color legend because it's the same as y axis
        axis.title = element_text(size = 12), # making all the text bigger
        axis.text = element_text(size = 11),
        strip.text.x = element_text(size = 12, face = "bold")) + # making facet titles bigger and bold
  scale_color_manual(values = c("#e37b7f", "#7b95d6", "#7de37e")) # set colors
```
