---
title: "Week 6 Homework"
author: "Micaela Chapuis"  
output: rmdformats::downcute
---

## Assignment Details

-   Take any data and create at least one table and one figure
-   Create a beautiful RMarkdown output → **Knit the file!**
-   Have headings with clear explanations
-   Use bold, italics, and lists at least once

```{r setup, include=FALSE}
# make sure all code shows (except this chunk), but no messages or warnings
# set the path for figure output
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../../Output/Homework/")
```

## Loading Libraries

```{r}
library(tidyverse)
library(here)
library(reactablefmtr)
```

## Loading Data
Using environmental data from Becker and Silbiger (2020)
```{r}
# Environmental data from each site
EnviroData<-read_csv(here("Week6","Data", "site.characteristics.data.csv")) %>% glimpse()

#data is in long format
```

## Preparing Data for Table and Figure

Taking means of each parameter by site and storing it in *site_param_means* because I **need** it for both my table and my figure
```{r}
site_param_means <- EnviroData %>%
  # group by species and take mean of each variable
  group_by(name, parameter.measured) %>%
  summarise(Mean = mean(values, na.rm = TRUE))
```


## Creating Table

Making a table of site data where we can *order the data in ascending or descending order*.  
The data is the average of each parameter selected (temperature, light, algal cover, coral cover, and CCA cover) by site (prepared above).  

Details:

-   Using the **{reactablefmtr}** package and the "nytimes" theme.

-   Details about the package are [here.](https://kcuilla.github.io/reactablefmtr/index.html)

-   The examples I followed are [here.](https://kcuilla.github.io/reactablefmtr/articles/themes.html)

```{r}
site_param_means %>%
  # pivot back to wide because I need the variables to be columns for my table
  pivot_wider(names_from = parameter.measured, #colums with the names for new column
              values_from = Mean) %>%
  select(name, temp, light, algal.cover, coral.cover, CCA.cover) %>% # select only the parameters I want + site name
  
  ## Creating table
  reactable(
    theme = nytimes(header_font_color = "black", # set theme for the table and specify color for header (column names)
                    font_color = "grey30"), # set font color to be a bit lighter
    
    # add column group header (Parameter) above the variables specified
    columnGroup = list(
      colGroup(name = "Parameter", columns = c("temp","light", "algal.cover", "coral.cover", "CCA.cover"))),
    
    # for each column: set the name and align it
    columns = list(
      name = colDef(name = "Site", # column name
                    align = "center"), # center column
      temp = colDef(name = "Temperature (°C)", # same as above
                    align = "center"),
      light = colDef(name = "Light", # same as above
                     align = "center"),
      algal.cover = colDef(name = "Algae Cover (%)", # same as above
                           align = "center"),
      coral.cover = colDef(name = "Coral Cover (%)", # same as above
                           align = "center"),
      CCA.cover = colDef(name = "CCA Cover (%)", # same as above
                         align = "center"))) %>%
  
  add_source(source = "Source: Becker and Silbiger (2020)", # add data source below table
             font_color = "#999999")
```

## Making a Figure

Making a *Lollipop Chart* ([example here](https://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html))

```{r w6_homework, fig.asp=1.1}
site_param_means %>%
  filter(parameter.measured %in% c("temp", "light", "algal.cover", "coral.cover", "CCA.cover")) %>% # using only the parameters I want to plot
  ggplot(aes(x=name, y=Mean, color = name)) + 
  facet_wrap(~parameter.measured, # facet by variable
             scales = "free", # let each variable have its own scale
             labeller = labeller(parameter.measured = c(`temp` = "Temperature (°C)", # set facet names
                                                        `light` = "Light",
                                                        `algal.cover` = "Algal Cover (%)",
                                                        `coral.cover` = "Coral Cover (%)",
                                                        `CCA.cover` = "CCA Cover (%)"))) +
  geom_point(size=5) + 
  geom_segment(aes(x=name, # geom segment draws lines between points (x,y) and (x end, y end)
                   xend=name,  # x and x end are both just site name because we don't want the line to move horizontally
                   y=0, 
                   yend=Mean), # vertically, line goes from the 0 until the value
               size = 1.5) + # making lines thicker
  labs(x = "Site", # axis labels
       y  = "Mean (Scales vary)") +
  theme_bw() +
  theme(legend.position = "none", # removing color legend because it's the same as y axis
        axis.title = element_text(size = 13), # making all the text bigger
        axis.text = element_text(size = 9),
        axis.text.x = element_text(angle=45, hjust=0.85), # adjusting site names so they can be read
        strip.text.x = element_text(size = 12, face = "bold")) + # making facet titles bigger and bold
  scale_color_manual(values = c("#00202e", "#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")) # set colors
```
