---
title: "w6a_lecture"
author: "Micaela Chapuis"
date: "2024-10-01"
output: html_document
---

## Notes

#### Metadata in YAML
* Always in between ---  
* Always goes in the top of the R Markdown document and includes metadata, style and type of output (type of format it will be saved as)  
* Always need title and output for it to work!  

Example:

--  
title: "title"    
output: 
  html_document:       (this spacing matters!)  
    toc: TRUE          (table of contents added)  
    toc_float: TRUE    (this makes the table of contents float, aka follow as you scroll)  
    theme: darkly      (we can change the theme, this one turns it into dark mode)  
--  


## Code Chunks 
Customizing chunk outputs

It can go inside each chunk { } 
```{r, inlcude = FALSE}
# this makes it so the code in this chunk specifically doesn't show up when knitted
```

Or it can be default for all chunks for the entire file
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # always show the code throughout the file
                      warning = FALSE) # no warnings show up throughout the file
```
^ This can be overriden by setting the customized output for a specific chunk later



