---
title: "w6b_lecture"
author: "Micaela Chapuis"
date: "2024-10-03"
output: html_document
---

# Week 6 - Online Lecture
## Working with RMarkdown

### Comments in R Markdown
<!-- This line is commented, won't show in the document. It needs to be closed -->
This line is not commented

### Working with Figures
We can change size, resolution, device (is it png, pdf, etc) and save in a specific folder.  
[Click here for details](https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing)

----------------------------------------------

### Practice! 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../../Output/")
```


#### Load Libraries

```{r}
library(palmerpenguins)
library(tidyverse)
library(kableExtra)
```

#### Making a plot!
```{r w6b_penguinplot, fig.asp=1.2, fig.align='center', fig.cap="This is a boxplot of penguins", dev="pdf"}
penguins %>% 
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot() + 
  geom_point(alpha = 0.3, position = "jitter") +
  theme_bw() +
  labs(y = "Body Mass (g)", 
       x = "Species")
```


**Changing Size Options:**   
- out.width="70%" (makes it 70% of the width)  
- fig.width=3 (makes de width 3 inches)  
- fig.height=4 (makes the height 4 inches)  
- fig.asp=1.2 (changes the aspect ratio to 1.2)  
- fig align="center"  



**Saving the Figure:**  
Set the file path for the figure using:  
- fig.path (can put this in a global option to send ALL figures there)  
- name your code chunk (this will be the figure name)  

Place between curly braces: {r -label-}

**IMPORTANT**: R automatically saves everything to the _file_ directory, NOT the _project_ directory. So we can't use the _here_ package.  
We have to give the file path. "../" means go up a directory. So "../Output/" means go out of the Scripts folder and to the output folder.  


We can also give it a caption with:  
- fig.cap="This is a bopxlot of penguins"  

To save the figure as a pdf instead of a png:  
- dev="pdf"

----------------------------------------------  
  
#### Tables
**We can type out tables by hand**  
The :'s on the second line indicates if it's left, center or right justified  
The number of dashes doesn't matter  


| Time   | Session  | Topic   |
|:-----  |:-------: |-------: |
| _left_ | _center_ | _right_ |
| 1:00   | 1        | Anatomy |
| 1:50   |          | *Break* |
| 2:00   | 2        | Tables  |
| 2:45   |          | *Break* |  


**Tables with Kable**   
Look up kable package to see all the ways we can edit our table  

```{r}
penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl()  %>% # make it a kable table
  kable_classic() %>% # change the format of the table
  row_spec(2, bold = T, color = "white", background = "red") %>% # makes the second row bold and gives it specific colors
  kable_styling(full_width = FALSE) # makes it narrower
```

**Other Table Packages**  
[Click here](https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/)  
Check out the _gt_ and _gtExtras_ packages  
There are ways to add plots or images to tables! 

----------------------------------------------  
  
#### RMarkdown Outputs

**Different Themes for Outputs**  
[Check out examples here!](https://www.datadreaming.org/posts/2018-04-11-r-markdown-theme-gallery/2018-04-11-r-markdown-theme-gallery)

**Output Options**
*github_document* makes a file that is easy to view on GitHub  
**NOTE**: ever file type has different limitations for the YAML  

[List of different types of outputs](https://rmarkdown.rstudio.com/lesson-9.html)


----------------------------------------------

#### YAML
The {ymlthis} package has a GUI for writing yml code  
Click *Tools* and then *Addins*, then scroll to "Write new RMarkdown or YAML file"