---
title: "week4a_homework"
author: "Micaela Chapuis"
date: "2024-09-18"
output: html_document
---

## Code Description
This script contains the code for Assignment 4a. 
There are 2 parts of this script: 
For part 1 we calculate the mean and variance of penguin body mass by species, island, and sex without any NAs.
For part 2 we filter out male penguins, then calculate the log body mass, then select only the columns for species, island, sex, and log body mass. Then we use these data to make a plot with clean and clear labels and that follows best practices. The plot is saved in the Week 4 Output folder.


## Loading libraries
```{r}
library(palmerpenguins)
library(tidyverse)
library(here)
```


## Loading in data
In this case, data comes from palmerpenguins package, so here we'll just look at it
```{r}
head(penguins)
```


## Analysis

### Part 1
Calculating mean and variance of penguin body mass, grouped by species, island and sex, and removing NAs
```{r}
data <- penguins %>%  #storing the results in a df called data
          drop_na(sex, island, species) %>%  # dropping NAs from all 3 columns just in case
          group_by(species, island, sex) %>% # grouping by species, island, and sex
          summarise(mean_mass = mean(body_mass_g, na.rm = TRUE), # calculating mean of body mass after grouping
                    mass_variance = var(body_mass_g, na.rm = TRUE)) # calculating variance of body mass after grouping

data # viewing our data to check if it worked
```

### Part 2
Filtering out male penguins, calculating the log body mass, selecting only the columns for species, island, sex, and log body mass. Then using these data to make a plot and save it in the Week 4 Output folder.
```{r}
penguins %>%
  filter(!(sex == "male")) %>%  # excluding just males (didn't say anything about excluding NAs or only including females)
  mutate(log_mass = log(body_mass_g)) %>% # calculating log body mass
  select(species, island, sex, log_mass) %>% # selecting only columns for species, island, sex, and log body mass
  ggplot(aes(x = island, y = log_mass, fill = species)) + # creating a plot using these data, colored by species
    geom_violin() + # adding violin plot geom
    labs(x = "Island", # setting labels
         y = "Log(Body mass (g))",
         fill = "Species") +
    scale_fill_viridis_d() + # setting color scale for the species fill
    theme_bw() + # setting a nice theme
    theme(axis.title = element_text(size = 14, face = "bold"), # manually making axis titles font size bigger + bold
        axis.text = element_text(size = 13, face = "italic"), #making axis text bigger + italic
        legend.title = element_text(size = 14, face = "bold"), # making legend title bigger + bold
        legend.text = element_text(size = 13), # making legend text bigger
        legend.position = "top") # moving the legend to the top

ggsave(here("Week4", "Output", "w4a_homework_plot.jpeg"), width = 7, height = 5)
```

