---
title: "week4_homework"
author: "Micaela Chapuis"
date: "2024-09-18"
output: html_document
---

## Code Description
This script contains the code for Assignment 4. 
There are 2 parts of this script: 
For part 1 we calculate the mean and variance of penguin body mass by species, island, and sex without any NAs.
For part 2 we filter out male penguins, then calculate the log body mass, then select only the columns for species, island, sex, and log body mass. Then we use these data to make a plot with clean and clear labels and that follows best practices. The plot is saved in the Week 4 Output folder.


## Loading libraries
```{r}
library(palmerpenguins)
library(tidyverse)
library(here)
```


##Loading in data
In this case, data comes from palmerpenguins package, so here we'll just look at it
```{r}
head(penguins)
```


## Analysis

### Part 1
```{r}
data <- penguins %>%  #storing the results in a df called data
          drop_na(sex) %>%  # dropping NAs only from sex because I know there's no NAs in species or island
          group_by(species, island, sex) %>% # grouping by species, island, and sex
          summarise(mean_mass = mean(body_mass_g, na.rm = TRUE), # calculating mean of body mass after grouping
                    mass_variance = var(body_mass_g, na.rm = TRUE)) # calculting variance of body mass after grouping

data # viewing our data to check if it worked
```

### Part 2
```{r}
penguins %>%
  filter(!(sex == "male")) %>%  # excluding males 
  mutate(log_mass = log(body_mass_g)) %>% # calculating log body mass
  select(species, island, sex, log_mass) %>% # selecting only columns for species, island, sex, and log body mass
  ggplot(aes(x = island, y = log_mass, fill = species)) + # creating a plot using these data
    geom_violin() + 
    labs(x = "Island",
         y = "Log(Body mass (g))",
         fill = "Sex") +
  theme_bw() +
  theme(axis.title = element_text(size = 12), # manually making all the text bigger
        axis.text = element_text(size = 11),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 11))
```

