---
title: "week4_lecture"
author: "Micaela Chapuis"
date: "2024-09-17"
output: html_document
---

# Week 4 Tuesday Lecture
## Plotting more penguin data!

Load libraries
```{r}
library(palmerpenguins)
library(tidyverse)
library(here)
```

Loading data (already loaded because it's from a package, but we'll look at it)
```{r}
glimpse(penguins)
```

## Analysis

### Filtering
First, look at data to make sure we know what our data looks like
```{r}
head(penguins)
```

Filter for only females
```{r}
filter(.data = penguins, sex == "female") 
```

Filter for penguins measured in 2008
```{r}
filter(penguins, year == 2008) # because it's numeric, we don't have to use quotes
```


Filter for penguins that have a body mass greater than 5000
```{r}
filter(penguins, body_mass_g > 5000)
```

Filter females greater than 5000 g
```{r}
filter(penguins, sex == "female", body_mass_g > 5000)
```

#### Boolean Operators
Same as above
```{r}
filter(penguins, sex == "female" & body_mass_g > 5000)
```

Now, using filter and boolean logic for:

Penguins collected in either 2008 or 2009
```{r}
filter(penguins, year == 2008 | year == 2009)
# filter(penguins, year %in% c("2008", "2009"))   # alternative option
```

Penguins that are not from the island Dream
```{r}
filter(penguins, !(island == "Dream"))
# filter(penguins, island != "Dream")  # "more common way" lol
```


Penguins in the species Adelie and Gentoo
```{r}
filter(penguins, species %in% c("Gentoo", "Adelie")) # using %in% is best practice when picking multiple things
```


### Mutate

```{r}
data <- mutate(penguins, 
               body_mass_kg = body_mass_g/1000, # convert mass to kg
               bill_length_depth = bill_length_mm/bill_depth_mm) # calculate ratio of bill length to depth

data
```

You can mutating multiple columns at once based on some criterion with:  
mutate_if(), mutate_at(), mutate_all()

link with examples in slides

#### ifelse()
Doing conditional tests with mutate()
ifelse() only allows for 2 things, there is a function called case_when() for multiple things
```{r}
data <- mutate(penguins,
               after_2008 = ifelse(year>2008, "After 2008", "Before 2008"))

# if year is greater than 2008, in the column write After 2008, otherwise write Before 2008
data
```


Try:
Create a column to add flipper length and body mass together
Use mutate and ifelse to create a new column where body mass greater than 4000 is labeled as big and everything else is small
```{r}
data2 <- mutate(penguins, 
                flipper_plus_mass = flipper_length_mm + body_mass_g,
                thicc_penguins = ifelse(body_mass_g>4000, "Big boi", "Smol boi"))
data2
```


### Pipe!!
pipe means "and then do"

```{r} 
penguins %>% # use penguins dataframe
  filter(sex == "female") %>%  #select female
  mutate(log_mass = log(body_mass_g)) #calculate log biomass
```

### Select
selecting certain columns to remain in the dataframe
```{r}
penguins %>%
  filter(sex == "female") %>%
  mutate(log_mass = log(body_mass_g)) %>%
  select(species, island, sex, log_mass)
```

Can also use select to rename our columns
```{r}
penguins %>%
  filter(sex == "female") %>%
  mutate(log_mass = log(body_mass_g)) %>%
  select(Species = species, island, sex, log_mass)
```

### Summarize

```{r}
penguins %>%
  summarise(mean_flipper = mean(flipper_length_mm, na.rm = TRUE))
```

```{r}
penguins %>%
  summarise(mean_flipper = mean(flipper_length_mm, na.rm = TRUE),
            min_flipper = min(flipper_length_mm, na.rm = TRUE))
```

### group_by
Summarizing values by certain groups
group_by() by itself doesn't do anything, but it is powerful when put before summarize.

```{r}
penguins %>%
  group_by(island, sex) %>%
  summarise(max_bill_length = max(bill_length_mm, na.rm = TRUE),
            mean_bill_length = mean(bill_length_mm, na.rm = TRUE))
```


### Removing NAs
drop_na() drops rows with NAs from a specific column

```{r}
penguins %>%
  drop_na(sex) %>% 
  group_by(island, sex) %>% 
  summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))
```

### Pipe into a ggplot
```{r}
penguins %>%
  drop_na(sex) %>%
  ggplot(aes(x = sex, y = flipper_length_mm)) +
    geom_boxplot()
```

