---
title: "week3_class_script"
author: "Micaela Chapuis"
date: "2024-09-10"
output: html_document
---
### Week 3 Lecture Script
This script was created during Week 3 Tuesday Lecture. 
In this script, we create a plot using penguin data from the Palmer Archipelago.

##### Reading in necessary libraries
```{r}
library(palmerpenguins)
library(tidyverse)
```

##### Looking at data
```{r}
view(penguins) #opens data in a separate tab
head(penguins) #gives first 6 lines
tail(penguins) #gives last 6 lines
glimpse(penguins) #shows what type of data is in each column and the first few
```

##### Making a plot
Going through the different steps of building up a plot
```{r}
ggplot(data = penguins) #this should give empty box because we're not mapping anything
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm)) # this one adds an x variable but because we don't have a y (or tell it how to                                            plot it) it's still an empty box
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm)) +   # added a y
  geom_point()    # this is the type of geom we want it to use
```

From here on, I will keep adding the rest of the code to this chunk
```{r}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     color = species)) + #this colors the points by species, has to go inside aes argument
  geom_point() +
  labs(title = "Bill depth and length",   # adding title label
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins", # subtitle label
       x = "Bill depth (mm)",  # x axis label
       y = "Bill length (mm)",  # y axis label
       color = "Species",  # legend label
       caption = "Source: Palmer Station LTER / palmerpenguins package") + #caption with data sources
  scale_color_viridis_d()   # adding a discrete color scale (designed to be preceived by viewers with common forms of    
                            #color blindness)
```

Now I make a new copy of the plot so I can practice messing with different aesthetic settings

```{r}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     color = species,
                     shape = island, # added different shapes for different islands 
                     size = body_mass_g, # use different size points for different penguin body masses
                     alpha = flipper_length_mm)) + # change transparency for different flipper lengths
  geom_point() +
  labs(title = "Bill depth and length",  
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", 
       y = "Bill length (mm)", 
       color = "Species", 
       shape = "Island", # added the legend to be Island for the different shapes
       size = "Body Mass (g)", # adding legend for different sizes
       alpha = "Flipper Length (mm)",  # adding legend for the different transparencies
       caption = "Source: Palmer Station LTER / palmerpenguins package") + 
  scale_color_viridis_d()  

#she ugly
```


#### Mapping vs Setting
In the previous chunk, size and alpha, etc were in the aesthetics (Mapping) so they were determined by variables
Here they're going to be a setting, so manually changed, not determined by variable values
```{r}
ggplot(penguins,
       aes(x=bill_depth_mm,
           y = bill_length_mm)) +
  geom_point(size = 2, alpha = 0.5)  # all the points will be size 2 and alpha 0.5 because it doesn't depend on what a variable's value is
```



#### Faceting
#####Facet_grid()
```{r}
ggplot(penguins,
       aes(x=bill_depth_mm,
           y=bill_length_mm)) +
  geom_point() +
  facet_grid(species~sex) #this will a grid with species as rows and sex as columns
# here we're not adding all the things that make it pretty (like labels) but we generally should
# grid will determine the dimensions for us
```

Now we flip rows and columns
```{r}
ggplot(penguins,
       aes(x=bill_depth_mm,
           y=bill_length_mm)) +
  geom_point() +
  facet_grid(sex~species) #this will a grid with sex as rows and species as columns
```
##### Facet_wrap 
```{r}
ggplot(penguins,
       aes(x=bill_depth_mm,
           y=bill_length_mm)) +
  geom_point() +
  facet_wrap(~ species) # this a ribbon where we determine the dimensions, grid will determine dimensions for us
```

Specify number of columns
```{r}
ggplot(penguins,
       aes(x=bill_depth_mm,
           y=bill_length_mm)) +
  geom_point() +
  facet_wrap(~ species, ncol = 2) #specify that we want to make it 2 columns 
```

##### Facet and color 
```{r}
ggplot(penguins,
       aes(x=bill_depth_mm,
           y=bill_length_mm,
           color = species)) +
  geom_point() +
  scale_color_viridis_d() + # up to here it's the same as before
  facet_grid(species~sex) + # we added the facet grid by species and sex 
  guides(color = FALSE) #by setting colors = false we delete the color legend because the facet already separates the colors by species
```
